\documentclass[11pt,letterpaper]{article}

% ── Packages ──────────────────────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{float}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

% ── Metadata ──────────────────────────────────────────────────────────────────
\title{%
  \textbf{AGI Pragma: A Decision Intelligence Core for Safe Agentic AI}%
}

\author{%
  Rafal Zabinski\\[4pt]
  Independent Research\\
  \texttt{github.com/zabinskirafal/AGI-Pragma-Core}
}

\date{}

% ── Code style ────────────────────────────────────────────────────────────────
\lstset{
  basicstyle=\ttfamily\small,
  frame=single,
  backgroundcolor=\color{gray!8},
  breaklines=true,
  captionpos=b,
}

% ══════════════════════════════════════════════════════════════════════════════
\begin{document}
\maketitle

% ── Abstract ──────────────────────────────────────────────────────────────────
\begin{abstract}
We present AGI Pragma, a seven-stage Decision Intelligence Core (DIC) that
enforces structured risk evaluation before agentic AI systems execute actions.
The DIC filters candidate actions through branching, Monte Carlo critical path
estimation, FMEA risk scoring, a decision gate, a four-state circuit breaker,
utility-based selection, and Bayesian belief update. We evaluate on three
benchmarks --- Snake, Maze, and Dynamic Threat Gridworld --- against two
baselines: a random policy and an A* pathfinder. The DIC matches A* on static
environments (100\% solve rate in Maze) and surpasses it where hazards move:
in Gridworld, A* with static obstacle avoidance solved 30/50 episodes while
the DIC's Monte Carlo forward simulation solved 39/50, a \textbf{$+$30\%
improvement over optimal pathfinding}. We further evaluate persistent episodic
memory and derive four empirical lessons about DIC calibration.
\end{abstract}

% ── 1. Introduction ───────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}

Agentic AI systems --- language models with tool access, autonomous planners,
multi-agent pipelines --- are increasingly capable of consequential real-world
action: executing code, modifying files, calling external APIs, interacting with
live infrastructure. Their failure modes differ qualitatively from those of
passive models. A hallucinated output is correctable; a dropped database table
or a committed security misconfiguration may not be.

Existing safety approaches operate at the wrong level of abstraction. Post-hoc
monitoring detects failures after execution. Constitutional AI and RLHF shape
preferences but provide no enforcement. Prompt engineering nudges behavior without
hard constraints. None of these mechanisms answers the question that matters in a
deployed agentic context: \textit{is this specific action safe to execute right
now, given the current state of the environment?}

The gap is not the absence of risk knowledge --- it is the absence of a structured
mechanism that converts risk knowledge into a binding pre-execution constraint.

We address this gap with the Decision Intelligence Core (DIC), a seven-stage
pipeline that sits between action generation and action execution. The DIC makes
no assumptions about the upstream generator: an LLM, a rule-based planner, or a
learned policy may sit above it. Its role is not to generate better actions but
to govern which generated actions are permitted to execute, and to produce an
auditable record of every decision.

This paper makes the following contributions:
\begin{enumerate}[leftmargin=*, nosep]
  \item A fully specified seven-stage DIC architecture with domain-agnostic
        safety properties and domain-specific utility extension points.
  \item Empirical validation across three environments with increasing risk
        complexity, against two baselines: a random policy and an A* pathfinder.
  \item Implementation and experimental evaluation of cross-session Bayesian
        episodic memory, with a characterisation of the conditions under which
        it provides measurable lift.
  \item Four empirically-derived lessons about DIC calibration applicable to
        any agentic deployment context.
\end{enumerate}

% ── 2. Method ─────────────────────────────────────────────────────────────────
\section{Method: The Decision Intelligence Core}
\label{sec:method}

The DIC processes one decision per time step. Given the current environment state
and a set of candidate actions, it produces a single approved action and a complete
audit trace. Figure~\ref{fig:architecture} shows the system position; the seven
stages are described below.

\begin{figure}[H]
\centering
\begin{verbatim}
  ┌─────────────────────────────┐
  │   Action Generator          │  LLM, planner, neural policy
  └────────────┬────────────────┘
               │ candidate actions
               ▼
  ┌─────────────────────────────┐
  │   Decision Intelligence     │  Branching → Critical Path → FMEA
  │   Core  (DIC)               │  → Gate → Circuit Breaker
  │                             │  → Selection → Belief Update
  └────────────┬────────────────┘
               │ approved action + audit trace
               ▼
  ┌─────────────────────────────┐
  │   Environment / Execution   │  real world, simulation, API
  └─────────────────────────────┘
\end{verbatim}
\caption{DIC position in the system. The governance layer is indifferent to
the upstream generator.}
\label{fig:architecture}
\end{figure}

\paragraph{Stage 1 --- Branching.}
Candidate actions are enumerated and any action leading to an immediately fatal
or physically impossible state is removed before evaluation. This stage is shared
with the random baseline in all experiments; measured performance improvements are
attributable to Stages 2--7 only.

\paragraph{Stage 2 --- Critical Path Estimation.}
Each surviving candidate is evaluated by Monte Carlo rollout. From the state
that would result from the candidate action, $N$ trajectories of depth $d$ are
simulated under a lightweight random policy on a cloned environment. Each
rollout records whether the trajectory terminates in failure or depth exhaustion
(treated as timeout). Per-candidate estimates are produced:
\begin{itemize}[nosep]
  \item $p_{\mathrm{death}}$: fraction of rollouts ending in catastrophic failure,
  \item $p_{\mathrm{trap}}$: fraction where $\geq 3$ of 4 neighbours are
        simultaneously occupied (irreversible state),
  \item $\mathbb{E}[\text{steps to failure}]$: mean steps until first terminal event.
\end{itemize}
All experiments use $N = 200$ rollouts and depth $d = 25$ (Snake) or $d = 50$
(Maze, Gridworld).

\paragraph{Stage 3 --- FMEA Risk Scoring.}
Each candidate receives a Risk Priority Number:
\begin{equation}
  \mathrm{RPN} = \mathrm{Severity} \times \mathrm{Occurrence} \times \mathrm{Detection}
              \times \mathrm{Reversibility}
  \label{eq:rpn}
\end{equation}
Severity captures consequence magnitude (range 1--10). Occurrence is derived from
the Monte Carlo estimates. Detection captures how silently a failure mode manifests
--- silent failures receive Detection $= 1$ (highest penalty). Reversibility
captures how easily an action's effects can be undone: $R = 0$ denotes a fully
reversible action (no lasting state change), $R = 10$ denotes a fully irreversible
one (e.g.\ permanent deletion, schema destruction). Each domain defines
its own FMEA table. The maximum RPN across all failure modes is used for gating.

\paragraph{Stage 4 --- Decision Integrity Gate.}
Any candidate whose maximum RPN meets or exceeds a configurable threshold $\tau$
is blocked before utility evaluation. $\tau = 240$ in all experiments. The gate
is a hard constraint; utility estimates cannot override it.

\paragraph{Stage 5 --- Circuit Breaker.}
The circuit breaker maintains a four-state autonomy level based on the current
maximum RPN:

\begin{table}[H]
\centering
\begin{tabular}{lll}
\toprule
\textbf{State} & \textbf{RPN range} & \textbf{Effect} \\
\midrule
OK   & $< 180$   & No restriction \\
WARN & 180--219  & Decision depth logged \\
SLOW & 220--259  & Decision depth restricted \\
STOP & $\geq 260$ & Action blocked \\
\bottomrule
\end{tabular}
\caption{Circuit breaker states. When all candidates are blocked the system falls
back to the least-RPN candidate rather than failing open.}
\label{tab:circuit_breaker}
\end{table}

\paragraph{Stage 6 --- Decision Selection.}
Among unblocked candidates, the agent selects by utility. The general form is:
\begin{equation}
  U(a) = (1 - p_{\mathrm{death}}) \cdot w_1
       + (1 - p_{\mathrm{trap}}) \cdot w_2
       - \mathrm{dist} \cdot w_3
       - \mathrm{revisits} \cdot w_4
       - \frac{\mathrm{RPN}}{1000}
  \label{eq:utility}
\end{equation}
Domain-specific terms are added where appropriate: a proximity penalty in
Gridworld; exact BFS path distance in Maze.

\paragraph{Stage 7 --- Belief Update.}
After each action, Bayesian Beta trackers update hazard rate estimates from
observed risk signals. The posterior after episode $N$ is available as the
prior for episode $N+1$ via persistent episodic memory
(Section~\ref{sec:memory}).

% ── 3. Experiments ────────────────────────────────────────────────────────────
\section{Experiments}
\label{sec:experiments}

\subsection{Environments}

\textbf{Snake} is a 10$\times$10 grid where the agent navigates a growing snake
body toward food. Self-collision and wall collision are terminal. The dominant
risk is trap: the snake's own body creates irreversible configurations.

\textbf{Maze} is a 15$\times$15 grid generated by recursive backtracking. Wall
hits are no-ops. The agent navigates from the top-left cell to the bottom-right.
A BFS distance table is precomputed from the goal at episode start; the utility
function uses exact topological path distance, not Manhattan distance.

\textbf{Dynamic Threat Gridworld} is a 15$\times$15 grid with five wandering
hazard agents that move randomly each step. Contact with a hazard from any
direction is terminal. WAIT is a first-class action. This is the first
environment in the suite where $p_{\mathrm{death}}$ varies meaningfully across
candidate actions at every step --- moving toward a hazard cluster has
genuinely higher $p_{\mathrm{death}}$ than WAIT or evasion.

\subsection{Baselines}

We use two baselines, each isolating a different layer of the DIC's contribution.

\textbf{Random baseline.} Selects uniformly at random from \texttt{safe\_actions()} ---
the same immediate-collision filter as Stage~1 of the DIC. Uses no pathfinding,
no FMEA, no Monte Carlo estimation. $\Delta(\text{A*}/\text{Random})$ measures the
value of pathfinding alone above shared action filtering.

\textbf{A* baseline.} Computes the shortest path to the goal via A* with Manhattan
distance heuristic, replanned every step. No DIC pipeline. Domain-specific
adaptations: Snake A* uses body-aware neighbour filtering (tail cell unblocked);
Gridworld A* treats current hazard positions as impassable and returns WAIT when
no clear path exists. $\Delta(\text{Pragma}/\text{A*})$ measures the value of
Monte Carlo forward simulation and FMEA above optimal static-obstacle pathfinding.

\subsection{Results}

All results are over 50 episodes per agent per benchmark, seeds 0--49.

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\toprule
\textbf{Metric} & \textbf{Random} & \textbf{A*} & \textbf{Pragma} & \textbf{$\Delta$(A*/Rnd)} & \textbf{$\Delta$(Pragma/A*)} \\
\midrule
Avg score  & 2.6   & 27.8  & 22.5  & $+$988\%  & $-$19\% \\
Max score  & 7     & 44    & 33    & $+$529\%  & $-$25\% \\
Avg steps  & 284.2 & 225.0 & 198.0 & $-$21\%   & $-$12\% \\
Avg reward & 8.6   & 128.8 & 101.4 & $+$1401\% & $-$21\% \\
\bottomrule
\end{tabular}
\caption{Snake --- 10$\times$10 grid, 50 episodes each. A* establishes a
performance ceiling; Pragma does not exceed it because DIC risk-avoidance
occasionally routes around the direct food path.}
\label{tab:snake}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\toprule
\textbf{Metric} & \textbf{Random} & \textbf{A*} & \textbf{Pragma} & \textbf{$\Delta$(A*/Rnd)} & \textbf{$\Delta$(Pragma/A*)} \\
\midrule
Solved             & 3/50 (6\%)  & 50/50 (100\%) & 50/50 (100\%) & $+$1567\% & 0\%     \\
Avg steps (all)    & 296.1       & 46.1          & 46.1          & $-$84\%   & 0\%     \\
Avg steps (solved) & 234.7       & 46.1          & 46.1          & $-$80\%   & 0\%     \\
Avg reward         & $-$38.3     & 5.5           & 5.5           & $+$114\%  & 0\%     \\
\bottomrule
\end{tabular}
\caption{Maze --- 15$\times$15 recursive backtracker, 50 episodes each. A* and
Pragma are equivalent on a static maze: BFS-guided A* finds the same path as the
DIC utility function using BFS distance. The DIC adds no overhead and no benefit.}
\label{tab:maze}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\toprule
\textbf{Metric} & \textbf{Random} & \textbf{A*} & \textbf{Pragma} & \textbf{$\Delta$(A*/Rnd)} & \textbf{$\Delta$(Pragma/A*)} \\
\midrule
Solved           & 0/50 (0\%)  & 30/50 (60\%) & 39/50 (78\%) & ---       & $+$30\% \\
Killed by hazard & 50/50       & 20/50        & 11/50        & $-$60\%   & $-$45\% \\
Timed out        & 0/50        & 0/50         & 0/50         & ---       & ---     \\
Avg steps        & 56.8        & 20.2         & 22.8         & $-$64\%   & $+$13\% \\
Avg reward       & $-$17.4     & 0.0          & 3.4          & $+$100\%  & ---     \\
\bottomrule
\end{tabular}
\caption{Dynamic Threat Gridworld --- 15$\times$15, 5 wandering hazards, 50
episodes each. \textbf{Key result}: A* with static hazard avoidance solved 30/50;
Pragma with Monte Carlo forward simulation solved 39/50 ($+$30\%). The DIC's
ability to project hazard trajectories forward in time is the source of the gain.}
\label{tab:gridworld}
\end{table}

\paragraph{Key finding.}
In Snake and Maze, A* establishes a performance ceiling that the DIC matches or
does not exceed. Gridworld is the decisive environment. A* treats hazard positions
as static obstacles and replans each step; it solved 30/50. The DIC projects hazard
movement forward via Monte Carlo rollouts, estimating $p_{\mathrm{death}}$ for
each candidate action under the actual hazard dynamics; it solved 39/50. The
\textbf{$+$30\% improvement over optimal static-obstacle pathfinding} is attributable
entirely to forward risk modeling. An agent that knows only where hazards are cannot
avoid a hazard that is about to enter its planned path; the DIC can.

% ── 4. Episodic Memory ────────────────────────────────────────────────────────
\section{Episodic Memory}
\label{sec:memory}

\subsection{Implementation}

Stage 7 produces Beta posterior parameters $(a, b)$ for each tracked hazard rate.
The episodic memory module persists this state to \texttt{artifacts/<benchmark>/memory.json}
at session end and loads it at session start, seeding Beta trackers with the
loaded $(a, b)$ rather than the uniform prior $\mathrm{Beta}(1, 1)$. An optional
decay factor $\alpha \in (0, 1]$ shrinks accumulated pseudo-counts toward the
uniform prior:
\begin{equation}
  a_{\mathrm{new}} = 1 + (a_{\mathrm{loaded}} - 1) \cdot \alpha
  \label{eq:decay}
\end{equation}

We additionally implemented $p_{\mathrm{death}}$ blending: at each step, the
Monte Carlo estimate is blended with the tracker's current mean:
\begin{equation}
  p_{\mathrm{death,adj}} = 0.7 \cdot p_{\mathrm{death}}^{\mathrm{MC}}
                         + 0.3 \cdot \mu_{\mathrm{tracker}}
  \label{eq:blend}
\end{equation}
The adjusted value propagates through FMEA, circuit breaker, utility, and the
subsequent belief update.

\subsection{Experiment and Results}

A two-pass comparison was run on all three benchmarks (50 episodes per pass,
seeds 0--49). Pass~1: no memory file present (uniform priors, cold start).
Pass~2: memory loaded from Pass~1 (warm start). Both prior seeding alone and
the blended $p_{\mathrm{death}}$ were tested.

\textbf{Result:} zero measurable difference across all metrics in all three
benchmarks under both implementations.

\subsection{Analysis}

Three structural reasons account for the null result.

\textbf{(i) Deterministic seeds remove stochasticity between passes.}
Both passes use \texttt{seed=0..49} on identical environment instances.
Monte Carlo rollouts use the episode seed as \texttt{seed\_base}, so
$p_{\mathrm{death}}^{\mathrm{MC}}$ is identical in both passes for every action
at every step. The blended value can only differ by the prior term, which is
itself derived from the same Pass~1 run.

\textbf{(ii) In-session accumulation floods the inter-session prior.}
Within 50 episodes, Beta trackers accumulate thousands of pseudo-counts
(e.g.\ $\mathrm{Beta}(10028, 1)$ for \texttt{death\_rate} in Snake). A loaded
prior of the same magnitude is immediately overwhelmed by in-session updates
converging to the same posterior. The prior's influence vanishes within the
first 3--4 episodes of Pass~2.

\textbf{(iii) Performance ceilings.} Maze achieves 50/50 and Snake averages
22.5 on deterministic seeds regardless of prior calibration. There is no
headroom for memory-enabled lift.

Episodic memory is an adaptation mechanism, not a performance mechanism. Its
value is proportional to: (a) stochasticity between sessions, (b) a cold-start
period where prior calibration can materially change early decisions, and
(c) available headroom below performance ceilings. These conditions are absent
in fixed-seed 50-episode benchmarks but are present in real deployments with
drifting environments and short session budgets.

% ── 5. Real-World Application ─────────────────────────────────────────────────
\section{Real-World Application}
\label{sec:application}

We implemented the DIC in two interactive demo systems and one REST service to
validate its behavior beyond grid-world simulation.
Each system instantiates the same seven-stage pipeline with a domain-specific
FMEA table, critical path analysis, and reversibility dimension
($\mathrm{RPN} = S \times O \times D \times R$, threshold $\tau = 2400$).

\subsection{File Operations Agent}

An LLM proposes filesystem actions (\textsc{read}, \textsc{write}, \textsc{delete})
inside a sandboxed directory. Stage~1 blocks any path that escapes the sandbox
or contains traversal components. FMEA failure modes are parameterised by
reversibility: write-new ($R=3$), write-overwrite ($R=8$), delete ($R=10$).
Occurrence is driven by the session-level Beta tracker.

In a five-step mock session the pipeline approved \textsc{write}~\texttt{plan.md}
(RPN~$420$, dominant mode \textit{unintended\_write}: $S{=}4$, $O{=}5$, $D{=}7$,
$R{=}3$) and blocked \textsc{delete}~\texttt{temp.txt} (RPN~$3150$, dominant mode
\textit{wrong\_file\_deleted}: $S{=}9$, $O{=}5$, $D{=}7$, $R{=}10$).
The blocking failure mode is not permanent data loss (detectable, $D{=}2$) but
\textit{path confusion}: an LLM may hallucinate the target filename and the error
is hard to catch before execution ($D{=}7$).

\subsection{Database Operations Agent}

An LLM proposes SQL operations (\textsc{select}, \textsc{insert}, \textsc{update},
\textsc{delete\_row}, \textsc{drop\_table}) against a SQLite database.
Stage~1 blocks system tables (\texttt{sqlite\_master} etc.)\ and rejects table
names containing injection-relevant characters.
Reversibility spans $R{=}1$ (\textsc{select}) to $R{=}10$ (\textsc{drop\_table}).

A four-step mock session approved \textsc{select} (RPN~$100$), \textsc{insert}
(RPN~$525$), and \textsc{update} with a \texttt{WHERE} condition
(RPN~$1512$, dominant mode \textit{wrong\_rows\_updated}: $S{=}7$, $O{=}6$,
$D{=}6$, $R{=}6$). \textsc{drop\_table} was blocked at RPN~$3780$ (dominant mode
\textit{cascade\_failure}: $S{=}9$, $O{=}7$, $D{=}6$, $R{=}10$).
Cascade failure scores higher than schema destruction because downstream
application references break silently ($D{=}6$) rather than immediately ($D{=}1$).
The database was preserved; the four-row \texttt{users} table remained intact.

\subsection{Controlled Experiment: DIC vs.\ No-DIC (50 Sessions)}

To quantify the firewall effect, we ran 50 sessions in each of two conditions
against the file-operations agent.
Every session proposed the identical five-action sequence:
\textsc{read}, \textsc{write}, \textsc{write}, \textsc{delete}, \textsc{done}.
The target file existed on disk before each session so the \textsc{delete}
was a real filesystem operation, not a no-op.

\begin{table}[H]
\centering
\begin{tabular}{lrr}
\toprule
\textbf{Metric} & \textbf{Without DIC} & \textbf{With DIC} \\
\midrule
Sessions run                        & 50          & 50 \\
Irreversible actions executed       & 50/50       & \textbf{0/50} \\
Irreversible actions blocked        & n/a         & 50/50 (100\%) \\
Files written                       & 100         & 100 \\
Task completion                     & 50/50 (100\%) & 50/50 (100\%) \\
\bottomrule
\end{tabular}
\caption{Controlled comparison: 50 sessions without DIC vs.\ 50 sessions with DIC
on the file-operations agent. Each session proposes one \textsc{delete} targeting
a real file.}
\label{tab:real_world}
\end{table}

DIC eliminated 100\% of irreversible operations while preserving full task
completion across all 50 sessions.
The dominant blocking failure mode was \textit{wrong\_file\_deleted}
($S{=}9$, $O{=}5$, $D{=}7$, $R{=}10$, RPN~$3150$) — LLM path hallucination
risk ($D{=}7$) rather than the more obvious permanent data loss ($D{=}2$).
Productive operations (\textsc{read}, \textsc{write}) were identical in both
conditions: the firewall constrains without obstructing.

\subsection{Decision Governance REST API}

The file-operations DIC is exposed as a stateful FastAPI service.
\texttt{POST /evaluate} accepts a \texttt{\{op, path, content, reason\}} payload
and returns \texttt{approved}, \texttt{block\_reason}, \texttt{max\_rpn},
\texttt{utility}, and the full seven-stage \texttt{stage\_log}.
The governor instance persists across requests; the Beta tracker and circuit
breaker accumulate session state identically to the interactive demo.
\texttt{GET /state} exposes the live session state; \texttt{POST /reset} returns
the governor to $\mathrm{Beta}(1,1)$.
The API pattern generalises directly to any domain with a DIC pipeline: replace
the FMEA table and critical path model, expose the same four routes.

% ── 6. Lessons Learned ────────────────────────────────────────────────────────
\section{Lessons Learned}
\label{sec:lessons}

\paragraph{L1 --- The safety pipeline is not the performance bottleneck.}
Across all three benchmarks and all calibration iterations, the FMEA gate and
circuit breaker operated correctly without modification. Every performance failure
was traceable to the utility function or critical path estimate. The pipeline is
a stable, domain-agnostic layer. Improving it did not improve performance;
improving the utility function and signal calibration did.

\paragraph{L2 --- The utility function is the primary design variable.}
Each benchmark improvement came from fixing the goal-progress signal, not the
safety constraints. Snake improved 57$\times$ by increasing the food-distance
weight from 0.2 to 1.5. Maze improved from 6\% to 100\% by replacing Manhattan
distance with BFS path distance. The correct architectural separation is
\textit{invariant safety} (DIC stages 1--5) and \textit{domain-specific progress
signal} (Stage~6 utility). Safety should remain fixed; the utility function is
tuned per domain.

\paragraph{L3 --- Critical path calibration must respect domain solution length.}
In Maze v1.0, rollout depth (25 steps) was shorter than the minimum maze solution
length under the rollout policy. $p_{\mathrm{death}}$ was never triggered,
silently producing zero for every candidate. The correct principle: calibrate
depth to the \textit{expected time-to-failure under the rollout policy}, not to
the episode step budget.

\paragraph{L4 --- Signal accuracy without variance is uninformative.}
After the Maze v1.0 calibration fix, $p_{\mathrm{death}} \approx 1.0$ for all
candidate actions --- accurate, but providing no basis for action selection. Useful
risk estimation requires variance across the action set, not merely correctness in
aggregate. The Gridworld environment is the first in this suite where
$p_{\mathrm{death}}$ is genuinely differentiating: moving toward a hazard cluster
has measurably higher $p_{\mathrm{death}}$ than WAIT or evasion.

% ── 7. Conclusion ─────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}

We presented the Decision Intelligence Core --- a seven-stage pre-execution
governance layer for agentic AI systems --- and validated it empirically across
three benchmark environments against two baselines: a random policy and an A*
pathfinder. The DIC achieved 100\% solve rate in a static maze (matching A*),
78\% solve rate in a dynamic hazard environment where random solved 0\% and A*
solved 60\%, and $+$988\% score improvement over random in Snake. The central
result is the Gridworld comparison: the DIC's Monte Carlo forward simulation
produced a \textbf{$+$30\% improvement over optimal static-obstacle pathfinding}
by reasoning about where hazards \textit{will be}, not just where they are.

The safety pipeline was stable across all conditions. Performance was governed by
utility function design and critical path calibration. Episodic memory via Bayesian
prior transfer is architecturally sound but requires stochastic, variable, or
short-session environments to produce measurable improvement over in-session
Bayesian updating alone.

The DIC enforces the distinction between \textit{intelligence that acts} and
\textit{intelligence that acts within justifiable bounds} --- making the latter a
property of the system rather than a property of any particular upstream model.

\bigskip
\noindent\textbf{Code and replication.}
All benchmark environments, agent implementations, and experiment runners are
available at \url{https://github.com/zabinskirafal/AGI-Pragma-Core}.

% ── End ───────────────────────────────────────────────────────────────────────
\end{document}
